---
title: "Using the WorldCat Search API"
# title: "using-the-worldcat-search-api"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the WorldCat Search API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# let's load this package before getting started
library(libbib)
```

_Intermediate knowledge of R is required to follow the examples here._

_Basic knowledge of how to manipulate `data.table`s if helpful but_
_following the examples should be possible without it._

## Introduction

`libbib`'s ability to communicate to the WorldCat Search API is probably
the most helpful capability of the package. As such, an entire vignette
dedicated to this powerful tool is warranted.

Also of note is that documentation on how to use this API
is scattered throughout the web, in different places, and
sometimes even no longer available and only accessible through the
internet archive. Because of this, the goal of this vignette is not
only to document `libbib`'s `worldcat_api_search` function, but to
provide a single location for compiling this archived documentation
and providing information/examples of what the WorldCat Search API is
capable, in general.

The terminology/nomenclature of the different WorldCat API offerings
is fuzzy and inconsistent. The specific kind of Search API that this package
offers usage of can, more specifically, be described as the
'WorldCat SRU Search API version 1'.

There is a version 2 of this API, but version 1 hasn't been sunsetted as
of yet.

First, let's speak of what this API is and what it is not.

## What the WorldCat SRU Search API is

This API is allows a developer to search for bibliographic records that
are cataloged in WorldCat.

The queries are made using [the SRU standard search protocol](https://en.wikipedia.org/wiki/Search/Retrieve_via_URL)
(Search/Retrieve via URL) using a standard query syntax called [CQL](https://en.wikipedia.org/wiki/Contextual_Query_Language)
(Contextual Query Language). The WorldCat Search API doesn't implement
all features of CQL; this vignette will illustrate SRU/CQL only
insofar as it is supported by this API.

This API is open to libraries that maintain both WorldCat Discovery and
OCLC Cataloging subscriptions and needs an API key (called a `WSKey`) to work.
A request can be made for a key (if your institution doesn't already have one)
via [this link.](https://help.oclc.org/Discovery_and_Reference/WorldCat_Discovery/Troubleshooting/How_do_I_request_a_WSKey_for_the_WorldCat_Search_API)

## What the WorldCat SRU Search API is not

This API is not the OpenSearch/Basic API, which doesn't allow field-specific
searches and only supports keyword-anywhere searches.

This is also not the same as using the
[advanced search option on worldcat.org](https://www.worldcat.org/advancedsearch).
That option only provides a subset of the bibliographic records in WorldCat and
is a far less powerful tool.

This API is most akin to using the "expert search" option in OCLC's
FirstSearch, using the WorldCat database, but differs in that (a) searching
the API is programmatically automate-able, and (b) the API allows for, still,
more powerful search queries.

## What `libbib`'s `worldcat_api_search` function provides

At its most basic, this function takes a SRU query and returns a
`data.table` with most of the bibliographic metadata from the MARCXml
that the API returns. We'll see how the behavior of this function can
be controlled by specifying certain function parameters.

This function also offers assistance with the SRU query syntax. Examples
of these aids will be explained later in the vignette.

## An example of usage and what the function returns

Let's look at an example of a simple query and what the returned
`data.table` looks like.

Specifically, we'll search for "Madame Bovary" by "Gustave Flaubert"
and we'll only show the first three results

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
library(libbib)      # load this package

result <- worldcat_api_search('$title="Madame Bovary" and
                                 $author="Gustave Flaubert"')

# get the column names
names(result)

# show the first three results
result[1:3,]
```

This should give you an idea of the rich information returned by the results
`data.table`. All of the information returned is from (or derived from)
the MARCXml that the API returns, save for

  - `total_wc_results`
    Which returns the number of the results the query yields in WorldCat even
    if the number of records/results requested is lower than this number.
    
  - `result_number`
    This number of the result, which is helpful if you're using a starting
    position other than `1` (the default)
    
  - `query`
    The final SRU query (after any of the SRU query assistance routines step
    in) that is sent to the API. This can be useful for debugging.

Hereafter, the output of the example queries will be abbreviated, truncated,
elided, or only show a subset of columns in order to save space and
aid in following the guide.

## What does an error look like

If you craft a query that yields no results, a message telling you such
will be displayed, and the return value is `NULL`.

If you make an error in the query syntax, no results will be returned, but
a diagnostic message returned from the server may tell you what went wrong.
Here are two examples...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, error=TRUE, eval=TRUE}
# missing ending double quotes
worldcat_api_search('$title="Madame Bovary and $langauge=greek')

# "$titley" is not a valid search index
worldcat_api_search('$titley="Madame Bovary" and $langauge=greek')

```


## `worldcat_api_search` by example / quick start

In guides like this, these is frequently tension between wanting to
be a complete references, and wanting to show cool/helpful examples,
at a glance.

In service of easing this tension, we'll first look at some examples
illustrating things this function can do and _then_ turn our attention
to an adequate explanation of boolean operators, relations operators,
search indexes, etc...

These examples are largely taken from the `examples` section of the
R documentation for this function.

These examples (and the ones further along in this vignette) will also
use `data.table` syntax to limit number of rows and columns returned
to aid reading.

We'll be focusing on the types of queries you can use in this section;
controlling the behavior of the function via changing parameters will
come later.

```{r, eval=FALSE}
library(libbib)

# title search for "The Brothers Karamazov"
results <- worldcat_api_search('$title="Brothers Karamazov"')

# Madame Bovary by Gustave Flaubert in Greek
sru <- '$author="Gustave Flaubert" and
          $title="Madame Bovary" and
          $language=greek'
results <- worldcat_api_search(sru)

# Hip Hop materials on wax, cassette, or CD published from
# 1987 to 1990
sru <- '(($material_type=cas or $material_type=cda or $material_type=lps)
           and $subject="Rap") and $year="1987-1990"'
results <- worldcat_api_search(sru)

# keyword search for "Common Lisp" for materials held
# at The New York Public Library
sru <- '$keyword="common lisp" and $holding_library=NYP'
results <- worldcat_api_search(sru)

# keyword search for "Common Lisp" for materials held
# at by any of the members of the "Manhattan Research
# Library Initiative" (MaRLI) joint borrowing program
# (New York Public Library, Columbia University, and
# New York University)
sru <- '($keyword="common lisp" and $holding_library=NYP)
          or ($keyword="common lisp" and $holding_library=ZYU)
          or ($keyword="common lisp" and $holding_library=ZCU)'
results <- worldcat_api_search(sru)

# Books (only books) about Ethics (by dewey division 170s or
# LC call number subject class "BJ") published in the 19th
# century
sru <- '($dewey="17*" or $lc_call="bj*") and $year="18*" and
           $material_type=bks'
results <- worldcat_api_search(sru)

# Materials on Musicology (by Dewey division 780s) at
# the New York Public Library and not held by any
# other insitution
sru <- '$dewey="78*" and $holding_library=NYP and
          $library_holdings_group=11'
results <- worldcat_api_search(sru)

# Search for materials on "Danger Music" published since 2010
results <- worldcat_api_search('$keyword="danger music" and $year="2010-"')
```


Now that we've seen these examples, sans explanation, we can now
have a closer look into the components of a query.
Broadly speaking, there are four concepts to be aware of...

  - relations operators
  
  - boolean operators
  
  - wildcards
  
  - search indexes

We'll be looking at each in this order, because I think that makes
the most sense. To illustrate the first three concepts, though,
we have to use search indexes, before a formal explanation of what
they are.

Briefly, a search index is a facet along which to search. In all of
the examples above, the search indexes were prefixed by a `$`
character (e.g. `$title`, `author`, `$keyword`, etc...)


## Relations operators

There are four relations operators available for use...

  - `=`
  
  - `exact`
  
  - `any`
  
  - `all`

It should be noted that not every relations operators is available
for use with every search index.

### `=`

This was the most common operator used in the quick-start examples above.
Though this operator can be most fully understood via constrast with the
next operator, `exact`, suffice it to say, for now, that using `=` means
that all of your search terms must match, without intervening words.
This is sometimes referred to as an "un-anchored" search.

If the phrase you're search for has spaces in it, you have to surround
it with double quotes. Since the SRU query to the `worldcat_api_search`
function must be a string, and strings can be made with single quotes
and double quotes, we need to surround the entire query with single
quotes.

We can get around this by "escaping" the quotes, using only single
quotes to surround the whole query is the most elegant approach.

You can use both spaces and `+` to separate the search index and the
search term on both sides of the relations operator...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
sru <- '$title+=+"Common Lisp"'
results <- worldcat_api_search(sru)
results[1:3, .(total_wc_results, title)]
```

is the same as...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
sru <- '$title = "Common Lisp"'
results <- worldcat_api_search(sru)
results[1:3, .(total_wc_results, title)]
```

But we'll be using spaces here.

The `=` relations operator is the only one that doesn't require any
space or `+` on either side of the operator, and we'll be using both
the spaced version and non-spaced version in these examples for this
operator, only.


### `exact` (anchored search)

The `exact` operator, in constrast with the `=` operator, signals to
the API that your search must match as an exact phrase, without
any other terms in the matching string. For this reason, this operator
is sometimes referred to as an "anchored search".

Here's an example using the same title search as above...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
results <- worldcat_api_search('$title exact "Common Lisp"')
results[1:3, .(total_wc_results, title)]
```

Note that the title "Practical Common Lisp" was returned by the
`=` operator query. The `exact` operator will not match this title
since it has another word/phrase in it besides "Common Lisp".

Using an `exact` search is probably a better fit than `=` for some
indexes and in some situations. For example, if you know the exact
title of a book (e.g. `$title exact "Brothers Karamazov")`, using
`exact`  will ensure that books with a title like _"A Guide to The_
_Brothers Karamazov"_ or _"The Brothers Karamazov in Pictures"_ are
not returned by the search.

That being said, we'll be using the `=` most heavily, and it's easier
to read.

### any

Using `any` means that any of your search terms (inside the double quotes)
can match. For example...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
results <- worldcat_api_search('$title any "Common Lisp"')
results[1:3, .(total_wc_results, title)]
```

matches the title "The book of common prayer" (decidedly _not_ a book
of the subject of Common Lisp) since it has the word
"common" in it. (Note that capitalization doesn't matter).

Using `any` in this context is tantamount to using an `or` boolean operator,
which we'll look at in the next section...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
results <- worldcat_api_search('$title = Common or $title = "Lisp"')
results[1:3, .(total_wc_results, title)]
```

(Note the same number and order of results)

### `all`

Using the `all` operator means that _all_ of the search terms must match but
the search terms can be in any order and have intervening terms in between.
For example, search like `$title all "Common Lisp"` can match a (fictional)
book book named _"Speaking with a lisp is common"_.

### A quick note before continuing

If your search term is one word, you do not have to use double quotes
to surround the term. For example, `$title = Ethics` and
`$title exact Ethics` word perfectly well.

If your search term as a single quote in it, it must be escaped so
that R doesn't interpret it as the end of the search string. That
being said, it appears as if you can just drop the single quote
and the search will carry on perfectly fine...

```{r, echo=FALSE}
Sys.sleep(1)
```

```{r, eval=TRUE}
results <- worldcat_api_search('$title exact "Finnegans Wake"')
results[1:3, .(total_wc_results, title, query)]

# yields the same results as

results <- worldcat_api_search('$title exact "Finnegan\'s Wake"')
results[1:3, .(total_wc_results, title, query)]
"
```

I suspect single quotes are automatically elided by the API.


## Boolean operators (and operator precedence)



## Wildcards


## Table of all available search indexes

  | Formal search index code | libbib alias |
  |--------------------------|--------------|
  | srw.kw | $keyword |
  | srw.ti | $title |
  | srw.ln | $language |
  | srw.au | $author |
  | srw.yr | $year |
  | srw.su | $subject |
  | srw.li | $holding_library |
  | srw.mt | $material_type |
  | srw.no | $oclc |
  | srw.lc | $lc_call |
  | srw.dd | $dewey |
  | srw.dn | $lccn |
  | srw.bn | $isbn |
  | srw.in | $issn |
  | srw.cg | $library_holdings_group |
  | srw.la | $language_code |
  | srw.pl | $place_of_publication |
  | srw.pb | $publisher |
  | srw.am | $access_method |
  | srw.cn | $corporate_conference_name |
  | srw.pc | $dlc_limit |
  | srw.dt | $document_type |
  | srw.gn | $government_document_number |
  | srw.mn | $music_publisher_number |
  | srw.nt | $notes |
  | srw.on | $open_digital_limit |
  | srw.pn | $personal_name |
  | srw.se | $series |
  | srw.sn | $standard_number |


## Selected Search Indexes in focus

### $holding_library / srw.

### $library_holdings_group / srw.

-----

## Combining this with the other API search functions in `libbib`

-----

## Sundry external / archived documentation links


- __WorldCat Search API current documentation__

  _scroll to SRU_
  
  Current documentation on the (SRU) WorldCat Search API call. Contains all
  index names and codes, parameters available, and the meaning of API
  status code return numbers
  
  <https://developer.api.oclc.org/wcv1>


- __WorldCat Search API > Using the API > Request Types > SRU__

  _(archived OCLC documentation link from 2013)_
  
  Examples of simple SRU requests and explanation of `frbrGrouping`
  and `servicelevel` parameters
  
  <https://web.archive.org/web/20130915101249/http://oclc.org/developer/documentation/worldcat-search-api/sru>


- __WorldCat Search API > Indexes__

  _(archived OCLC documentation link from 2013)_
  
  Explation of relations operators, boolean operators, title index
  stop-words (removed words), the MARC subfields search by (only) the subject
  index, and explation of the wildcard characters `*`; `#`; and `?`.
  
  <https://web.archive.org/web/20130530034745/http://oclc.org/developer/documentation/worldcat-search-api/indexes>


- __WorldCat Search API > Indexes > Complete List of Indexes__

  _(archived OCLC documentation link from 2013)_
  
  Explation of relations operators and a complete list of all indexes, their
  respective SRU index code, the relations available for use with each, and
  the MARC fields search for each index. Here, the SRU index codes have a
  prefix of `sru`, but it's really `srw`.

  <https://web.archive.org/web/20130915101232/http://oclc.org/developer/documentation/worldcat-search-api/complete-list-indexes>


- __Material Types Names and Codes__

  A complete list of Material Types you can search for, their codes 
  
  For use with the `srw.mt`/`$material_type` index.

  [Material Type names and codes - OCLC Support](https://help.oclc.org/Librarian_Toolbox/Searching_WorldCat_Indexes/Bibliographic_records/Material_Type_names_and_codes?sl=en)
  
  [Archived from original for posterity](https://web.archive.org/web/20210425140401/https://help.oclc.org/Librarian_Toolbox/Searching_WorldCat_Indexes/Bibliographic_records/Material_Type_names_and_codes?sl=en>)


- __WorldCat Search API > Indexes > Tips for specific indexes__

  _(archived OCLC documentation link from 2013)_
  
  Contains a lot of very helpful information about some of the different
  search indexes, the stop words they use, normalization rules, and
  a very helpful cross-walk on the "Number of Holding Libraries" index.
  
  <https://web.archive.org/web/20120311003221/http://oclc.org/developer/documentation/worldcat-search-api/tips-specific-indexes>


- __WorldCat Search API > Using the API > Parameters__

  _(archived OCLC documentation link from 2013)_

  Helpful information about what the different SRU search parameters mean.
  Remember that the parameters of the `worldcat_api_search` function (a) only
  contain a subset of these, and (b) may have slightly different names
  <https://web.archive.org/web/20121101034843/http://oclc.org/developer/documentation/worldcat-search-api/parameters>
  
  
- __Finding Institution Codes__

  A search box for searching institution names and returning instituion
  codes suitable for use in the holding institution search index.
  
  <https://www.oclc.org/en/contacts/libraries.html>


- __Searching WorldCat Indexes__

  _(archived OCLC documentation link from 2012)_
  
  Information about search indexes that are largely available in the above
  links. Not related to the (SRU) Search API but could be helpful, anyway.
  
  <https://web.archive.org/web/20120509072600/http://www.oclc.org/us/en/support/documentation/worldcat/searching/searchworldcatindexes/default.htm>
  
  
- __CQL specification__

  Helpful information about CQL query syntax. Not all is applicable to
  the WorldCat Search API that is discussed here.

  <https://www.loc.gov/standards/sru/cql/spec.html>


- __Searching WorldCat Indexes__

  Current page with links containing a lot of information about
  wielding WorldCat searches. Not all information is application to
  the WorldCat Search API that is discussed here, and most of the
  most helpful information is already covered in the links above.

  <https://help.oclc.org/Librarian_Toolbox/Searching_WorldCat_Indexes>



## Similar projects / code

- __A Ruby Gem to communicate with the WorldCat Search API__

  <https://github.com/vivien/worldcat>


- __Python scripts using the WorldCat search API__

  <https://github.com/ysugiyama3/worldcat-search-api>


## Papers/Articles

- __Using the WorldCat API to Develop Data-Driven Decision-Making for Gifts-in-Kind__

  <https://journals.ala.org/index.php/lrts/article/view/7074/9638>


- __Destroyer and Preserver, Hear, Oh Hear! Not All Uncirculated Books Must Chariotest to a Dark Wintry Bed: How We Used the Books Must Chariotest to a Dark Wintry Bed: How We Used the OCLC WorldCat Search API to Inform Our Weeding Decisions with OCLC WorldCat Search API to Inform Our Weeding Decisions with Holdings Data Holdings Data__

  <https://docs.lib.purdue.edu/cgi/viewcontent.cgi?article=2026&context=charleston>
